Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2024-08-26 Thesis/04-chap4.qmd
add caption for every plot in chap 4

To get started quickly, only calls two functions are necessary to obtain a summary of the automated residual assessment:

@@

1. Load the package using the `library()` function.
2. Construct a checker with two inputs: a linear regression model and a pre-trained Keras model [@chollet2015keras].
3. Call the `check()` method of the checker with default arguments. This predicts the VSS for the true residual plot, 100 null plots, and 100 bootstrapped plots, storing the predictions internally. A concise report of the check results will be printed.

@@

The method returns a `ggplot` object, which can be saved as a PNG file in the following module. If no data is provided, the method will use `get_fitted_and_resid()` to retrieve the fitted values and residuals from the regression model stored in the checker.

```{r echo=TRUE, fig.height=3, fig.width=4}
checker$plot_resid()
```

@@

```{r echo = TRUE, fig.height=3, fig.width=4}
checker$null_method() |>
  checker$plot_resid()
```

@@

After executing the `check()` method, `autovi` offers two visualization options for the assessment result through the `summary_plot()` method, including the density plot and the rank plot. We have already discussed and interpreted the density plot in @sec-autovi-usage. Here, we would like to highlight the flexibility in choosing which elements to display in the density plot. For instance, you can omit the bootstrapped distribution by setting `boot_dist` to `NULL`. Similarly, you can hide the null distribution (`null_dist`), the $p$-value (`p_value`), or the likelihood ratio (`likelihood_ratio`) as needed. The following example demonstrates how to create a summary plot without the results from bootstrapped plots.

```{r echo=TRUE}
checker$summary_plot(boot_dist = NULL,
                     likelihood_ratio = NULL)
```

@@

The rank plot, creating by setting `type` to "rank", is a bar plot where the $x$-axis represents the rank and the $y$-axis shows the VSS. The bar that is colored in red correspoding to the VSS of the true residual plot. By examining the rank plot, you can intuitively understand how the observed VSS compares to the null VSSs and identify any outliers in the null distribution.

```{r echo=TRUE}
checker$summary_plot(type = "rank")
```

@@

<!-- In @sec-autovi-implementation, we discussed how `autovi` relies on several Python libraries, with a particularly strong dependency on `TensorFlow`. Managing a Python environment and correctly installing `TensorFlow` on a local machine can be challenging for many users. Moreover, `TensorFlow` is a massive library that undergoes continuous development, which inevitably leads to compatibility issues arising from differences in library versions. These challenges can create significant barriers for users who want to perform residual assessments with the `autovi` package.

Recognizing these potential barriers, we were motivated to design and implement a web interface called `autovi.web`. This web-based solution offers several major advantages. First, it eliminates dependency issues, so users no longer need to struggle with complex Python environments or worry about installing and maintaining the correct versions of libraries. The web interface handles all these dependencies on the server side. Second, `autovi.web` lowers the entry barrier by being user-friendly and accessible to individuals who may not be familiar with R programming. This broadens the potential user base of `autovi`, allowing more researchers and analysts to benefit from its capabilities. Third, the web interface can be updated centrally, ensuring that all users always have access to the latest features and improvements without needing to manage updates locally. Lastly, `autovi.web` offers cross-platform accessibility, allowing users to access it from any device with a web browser, increasing flexibility and convenience.

`autovi.web` is available at [autoviweb.netlify.app](autoviweb.netlify.app). The implementation discussed in this chapter is based on `autovi.web` version 0.1.0. By providing this web interface, we aim to significantly reduce the technical hurdles associated with using `autovi`, making advanced residual assessment techniques more accessible to a wider audience of researchers and data analysts. This approach aligns with modern trends in data science tools, where web-based interfaces are increasingly used to make advanced analytical techniques more widely available.
-->

@@

Determining the best way to implement the interface was difficult. In our initial planning for `autovi.web`, we considered implementing the entire web application using the `webr` framework [@webr], which would have allowed the entire application to run directly in the user's browser. However, this approach was not feasible at the time of writing this chapter. The reason is that one of the R packages `autovi` depends on the R package `splancs` [@splancs], which uses compiled Fortran code. A working Emscripten (XXX reference?) version of this package, which would be required for `webr`, was not available.

@@

While the R package `autovi` aims to provide tools that can be extended to broader visual inference applications, `autovi.web` is only focus on to for providing  a straightforward and clean user interface. An overview of the graphical user interface of `autovi.web` is provided in @fig-autovi-web. This is the default view of the web application, and there are five regions that user can mainly interact with. Region 1 of @fig-autovi-web is a sidebar menu which can switch between the analysis page and the information page. The analysis page is the focus of this section.

@@

The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in each panel shown in @fig-autovi-web. Thre are two example datasets provided by the web application, as mentioned in @sec-autovi-web-design. The single residual plot example uses the `dino` dataset from the R package `datasauRus` [@datasaurus]. The lineup example uses residuals from a simulated regression model that has a non-linearity issue. We will walk through the lineup example to further demonstrate the workflow of the web application.

>>>>

To get started quickly, only calls three functions are necessary to obtain a summary of the automated residual assessment:

@@

1. Load the package using the `library()` function.
2. Construct a checker with a linear regression model.
3. Call the `check()` method of the checker with default arguments. This predicts the VSS for the true residual plot, 100 null plots, and 100 bootstrapped plots, storing the predictions internally. A concise report of the check results will be printed.

@@

The method returns a `ggplot` object (@fig-autovi-plot-resid), which can be saved as a PNG file in the following module. If no data is provided, the method will use `get_fitted_and_resid()` to retrieve the fitted values and residuals from the regression model stored in the checker.

```{r fig-autovi-plot-resid, echo=TRUE, fig.height=3, fig.width=4, fig.cap = "Residual plot of the regression model stored in the checker."}
checker$plot_resid()
```

@@

@@

After executing the `check()` method, `autovi` offers two visualization options for the assessment result through the `summary_plot()` method, including the density plot and the rank plot. We have already discussed and interpreted the density plot in @sec-autovi-usage. Here, we would like to highlight the flexibility in choosing which elements to display in the density plot as shown in @fig-autovi-summary-plot-option. For instance, you can omit the bootstrapped distribution by setting `boot_dist` to `NULL`. Similarly, you can hide the null distribution (`null_dist`), the $p$-value (`p_value`), or the likelihood ratio (`likelihood_ratio`) as needed. The following example demonstrates how to create a summary plot without the results from bootstrapped plots.

```{r fig-autovi-summary-plot-option, echo=TRUE, fig.cap = "Density plot of the VSS for null plots."}
checker$summary_plot(boot_dist = NULL,
                     likelihood_ratio = NULL)
```

@@

The rank plot (@fig-autovi-rank-plot), creating by setting `type` to "rank", is a bar plot where the $x$-axis represents the rank and the $y$-axis shows the VSS. The bar that is coloured in red corresponding to the VSS of the true residual plot. By examining the rank plot, you can intuitively understand how the observed VSS compares to the null VSSs and identify any outliers in the null distribution.

```{r fig-autovi-rank-plot, echo=TRUE, fig.cap = "Density plot of the VSS for null plots."}
checker$summary_plot(type = "rank")
```

@@

The web interface `autovi.web` is available at [autoviweb.netlify.app](autoviweb.netlify.app). The implementation discussed in this chapter is based on `autovi.web` version 0.1.0.

@@

Determining the best way to implement the interface was difficult. In our initial planning for `autovi.web`, we considered implementing the entire web application using the `webr` framework [@webr], which would have allowed the entire application to run directly in the user's browser. However, this approach was not feasible at the time of writing this chapter. The reason is that one of the R packages `autovi` depends on the R package `splancs` [@splancs], which uses compiled Fortran code. A working Emscripten [@zakai2011emscripten] version of this package, which would be required for `webr`, was not available.

@@

While the R package `autovi` aims to provide tools that can be extended to broader visual inference applications, `autovi.web` is only focus on providing  a straightforward and clean user interface. An overview of the graphical user interface of `autovi.web` is provided in @fig-autovi-web. This is the default view of the web application, and there are five regions that user can mainly interact with. Region 1 of @fig-autovi-web is a sidebar menu which can switch between the analysis page and the information page. The analysis page is the focus of this section.

@@

The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in each panel shown in @fig-autovi-web. There are two example datasets provided by the web application, as mentioned in @sec-autovi-web-design. The single residual plot example uses the `dino` dataset from the R package `datasauRus` [@datasaurus]. The lineup example uses residuals from a simulated regression model that has a non-linearity issue. We will walk through the lineup example to further demonstrate the workflow of the web application.

### 2024-08-26 Thesis/C-appC.qmd
add caption for every plot in chap 4

```{r}
EXT_AUTO_VI <- bandicoot::new_class(AUTO_VI, class_name = "EXT_AUTO_VI")
print(EXT_AUTO_VI)
```

@@

```{r}
names(EXT_AUTO_VI)
```

@@

```{r}
bandicoot::register_method(
  EXT_AUTO_VI, 
  null_method = function(fitted_model = self$fitted_model) {
    data <- self$get_fitted_and_resid(fitted_model)
    residual_sd <- sd(data$.resid)
    data$.resid <- rnorm(nrow(data),
                         sd = residual_sd)
    return(data)
  }
)

@@

```{r}
ext_auto_vi <- function(fitted_model,
                        keras_model = NULL,
                        data = NULL,
                        node_index = 1L,
                        env = new.env(parent = parent.frame()),
                        init_call = sys.call()) {
  EXT_AUTO_VI$instantiate(fitted_model = fitted_model,
                          keras_model = keras_model,
                          data = data,
                          node_index = node_index,
                          env = env,
                          init_call = init_call)
}

>>>>

```{r echo=TRUE}
library(autovi)
EXT_AUTO_VI <- bandicoot::new_class(AUTO_VI, class_name = "EXT_AUTO_VI")
print(EXT_AUTO_VI)
```

@@

```{r echo=TRUE}
names(EXT_AUTO_VI)
```

@@

```{r echo=TRUE}
bandicoot::register_method(
  EXT_AUTO_VI, 
  null_method = function(fitted_model = self$fitted_model) {
    data <- self$get_fitted_and_resid(fitted_model)
    residual_sd <- sd(data$.resid)
    data$.resid <- rnorm(nrow(data),
                         sd = residual_sd)
    return(data)
  }
)

@@

```{r echo=TRUE}
ext_auto_vi <- function(fitted_model,
                        keras_model = NULL,
                        data = NULL,
                        node_index = 1L,
                        env = new.env(parent = parent.frame()),
                        init_call = sys.call()) {
  EXT_AUTO_VI$instantiate(fitted_model = fitted_model,
                          keras_model = keras_model,
                          data = data,
                          node_index = node_index,
                          env = env,
                          init_call = init_call)
}

### 2024-08-26 Thesis/abstract.qmd
add caption for every plot in chap 4

This research presents three original contributions. The first contribution ([Chapter 2](#sec-first-paper)) provides evidence for the effectiveness of visual inference in regression diagnostics through a human subject experiment, demonstrating the benefits of using the lineup protocol for reliable and consistent reading of residual plots. The second contribution ([Chapter 3](#sec-second-paper)) introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitation of the lineup protocol. The third contribution ([Chapter 4](#sec-third-paper)) presents an R package and Shiny app, providing a user-friendly interface for analysts to leverage the computer vision model and supporting tools for diagnostic purposes. These contributions advance the field of artificial intelligence for data visualization, enabling more efficient and accurate regression diagnostics.

>>>>

This research presents three original contributions. The first contribution Chapter 2 provides evidence for the effectiveness of visual inference in regression diagnostics through a human subject experiment, demonstrating the benefits of using the lineup protocol for reliable and consistent reading of residual plots. The second contribution Chapter 3 introduces a computer vision model to automate the assessment of residual plots, addressing the scalability limitation of the lineup protocol. The third contribution Chapter 4 presents an R package and Shiny app, providing a user-friendly interface for analysts to leverage the computer vision model and supporting tools for diagnostic purposes. These contributions advance the field of artificial intelligence for data visualization, enabling more efficient and accurate regression diagnostics.
